---
output:
  pdf_document:
    includes:
      in_header: preamble.tex
      before_body: titlepage.tex
    keep_tex:  true
    extra_dependencies: ["booktabs", "threeparttable"]
bibliography: bibliography.bib
link-citations: true
geometry: "left=2cm,right=3cm,top=3cm,bottom=3cm"
papersize: a4
fontsize: 12pt
indent: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(gtools)
load("results.RData")
round_ <- function(x){round(x, digits = 4)}
options(scipen = 6)
```

\tableofcontents
\newpage
\onehalfspacing

# Introduction

Climate change, its consequences, and the necessity to limit them are arguably the most important political challenge facing governments in the twenty-first century. According to the latest estimates compiled by the intergovernmental panel on climate change (IPCC), the consequences of global warming exceeding 1.5°C or even 2°C appear to include vast increases in the frequency and severity of extreme weather events such as floods, heat waves, and fire weather [@ipcc2021a 23], and these effects are irreversible for centuries or even millenia (Ibid., 28). To minimize these threats, human CO$_2$ emissions must be reduced to net zero around 2050 [@ipcc2018 12]. This goal of carbon neutrality can only be reached through ambitious and economically significant policy measures, which are often politically divisive. In democracies, citizens decide, either directly or indirectly, whether and when such climate policies are adopted. Thus, the attitudes of these citizens towards climate protection in general, as well as towards individual policy projects, such as carbon taxation and measures towards energy transition, are essential to the global fight against climate change. Even though over 90% of citizens in most countries now accept human-made climate change as a reality [@charlton2019], specific policies designed to address it are far more contested. Examples such as the "yellow vests" protest wave in France show that while many citizens may agree with the general problem statement, significant measures such as increased taxation of goods with a high carbon footprint are not guaranteed to enjoy widespread support.

Previous works addressing individual-level attitudes towards climate change and environmental policies have focused on the influence of socio-demographic factors such as age, gender, income and level of education [@patchen2006; @douenne2020; @colvin2021]. Others have investigated individuals' perception of the risks and fairness of specific policies [cf. @drews2016 6-8], as well as the relationship between general political orientations (political trust, ideological leanings, party orientation) and various expressions of political environmentalism [@patchen2006; @dalton2009; @dunlap2001; @colvin2021]. While all of these variables can to some extent explain variations in the level of support for climate policies, they currently only offer partial explanations of the phenomenon, and there are still many missing puzzle pieces. With the additional urgency that the climate issue has received in the recent past, it seems relevant to investigate other potential sources of variance. One such source might be an inherent effect of different individual personalities. In trait psychology, it has become widely accepted that personality traits can be structured into five basic factors, often referred to as the "Big Five" [@goldberg1990; @mccrae1987; @mccrae2003]. These factors (Extraversion, Agreeableness, Conscientiousness, Neuroticism, and Openness to Experience) have already been shown to predict individual-level environmentalism or environmentally-significant behaviours [@hirsh2010; @milfont2012; @brick2016]. Recently, the widespread adoption of the Big Five model in political science following @mondak2010 has led to a surge in publications linking personality traits to various political attitudes and behaviours [cf. @gerber2011 for an overview; more recent examples include @bakker2016; @fatke2019; @aichholzer2021].

This paper aims to connect the two above-mentioned applications of the Big Five model by relating it to political environmentalism, which is conceptualized as an individual's support for ambitious political measures against climate change, and their tendency to prioritize environmental over economic concerns in the political sphere. How do basic personality traits influence individual support for ambitious climate policy? To answer that question, I pursue a two-step approach. First, I review previous findings in the fields of trait psychology, political psychology, and political sociology about the determinants of political environmentalism and the relationship between the Big Five and political attitudes, as well as environmentalism and environmentally significant behaviours. Since my research interest in the later chapters is mostly deductive, I am particularly interested in theoretical arguments and causal hypotheses. Based on this literature review, I formulate and test a set of hypotheses regarding the influence of the individual personality traits on four dependent variables: One general measure of the degree to which individuals prioritize climate protection over economic concerns, and three measures of support for specific climate-related policies (A tax on CO$_2$ emissions, a ban on new registration of combustion engine cars after 2030, and the continued use of nuclear power as an energy source). In addition to the Big Five model, I include an additional sixth trait construct, measuring an individual's tendency to embrace or eschew risk. I analyse survey data from the German Longitudinal Election Study 2016-2020 panel [@gles2021] to test the hypotheses empirically, using four multivariate OLS regression models. Results indicate that... !!INSERT!!

The paper contributes to the psychological findings on the antecedents of environmentalist attitudes and environmentally significant behavior by adding an explicitly political dimension (support for far-reaching policies, rather than micro-level behaviours of private citizens). In the field of political science, it adds to a growing body of research regarding the influence of citizens' innate personality traits on their political attitudes, values, and behaviours. Furthermore, insights into the personality traits of political environmentalists or anti-environmentalists may help policy makers better understand how to successfully communicate effective climate policies to voters and gain support for such endeavors.

# 1 Literature Review

## 1.1 Public support for climate policies

It is universally accepted almost as a truism that "public opinion is a key determinant of policy change in democratic countries" [@drews2016 2]. Consequently, the question of how individual citizens, who together constitute 'the public', form opinions on political issues and how those opinions then affect political practice constitutes one of the most-studied sub-fields of political science. Theorists and empirical researchers alike have studied it using a large number of more or less overlapping terms, including public opinion, public support, political support, political attitudes, and many others. In this chapter, I will focus primarily on the concept of public support, but will occasionally also refer to neighboring concepts and terminologies.

Probably the most influential conceptualization of the term "political support" was developed by David @easton1975. Easton introduces two key distinctions: Firstly, he distinguishes between support as "overt behavior, such as action or advocacy" [@easton1975 436] versus support as a positive attitude or favorable disposition towards a given object, which may be a person, institution, idea, policy, or any other political entity. This distinction might be referred to as *passive* and *active* support, though Easton does not use this terminology himself. Secondly, Easton divides political support into *specific* and *diffuse* support. Specific support refers to satisfaction with an individual's "conditions of life" (Ibid.), the incumbent political authorities and their actions, or certain political entities or policies. In contrast, diffuse support is concerned with "the underlying order of life" (Ibid.), i.e. the political and economic system as a whole. As Easton points out, there are frequently situations in which diffuse support does not falter even in the absence of specific support for the incumbent authorities. Within Easton's framework, the kind of support this paper deals with is clearly located on the side of specific support, since I am interested in citizens' support regarding a specific political field, climate policy, and even specific political projects within that field. In academic literature, this type of support is more often referred to as "public support" [e.g. by @inglehart1995; @drews2016; @vanboven2018], while the term "political support" is more often used when discussing support for political institutions, authorities, and office-holders [e.g. by @dalton1998]. 

Regarding the distinction between active and passive support, the lines are slightly more blurred. Obviously, support only becomes politically consequential (and thus interesting from a behavioral political science perspective) when it motivates or includes political behavior. Nevertheless, it seems obvious that, in order for such active support to manifest, passive support is a necessary condition. As Easton puts it: "One person is unlikely to support another in the senses mentioned here unless at the same time [they are] favorably disposed towards [them]" [@easton1975 436]. It is thus a good idea to first investigate the preconditions of positive attitudes towards climate policies, narrowing down the field of possibilities, and then concern oneself with the conditions under which this passive support translates into supportive behavior such as advocacy or activism. In this paper, I will attempt to do the former, and leave the latter for further research. However, it is important to note that this relatively narrow focus on attitudes rather than behaviours is externally motivated by constraints on the length of the paper, and should not be misinterpreted as a theoretical marginalization of active or behavioral support as an object of study. The kind of support that I am investigating, namely attitudinal support for effective political measures against climate change, broadly overlaps with the second of the four kinds of environmentally significant behavior identified by Stern, "nonactivist behavior in the public sphere" [@stern2000 409]. Stern's usage of the word "behavior" may be somewhat confusing here, but makes sense when one considers that he contrasts this nonactivist behavior, by which me means the public expression of supportive attitudes, with "environmental activism" (Ibid.). For Stern, expressing a supportive attitude towards environmental policy constitutes a behavior simply because it may influence public policy (Ibid.).

One attempt at a comprehensive model explaining "public attitudes and behavior about climate change" was made by @patchen2006. Patchen considers environmentally relevant behavior (which includes the public expression of policy-related attitudes) to be immediately determined by a person's "emotions about the conditions of the environment", along with a cost-benefit analysis and the person's sense of ability or self-efficacy [@patchen2006 4]. According to Patchen, these intermediary variables are themselves caused by a set of underlying determinants, which he divides into two broad categories: "social influences" and "personal characteristics" (Ibid.). Social influences and include demographic characteristics such as age, gender, social class and religious background, as well as exogenous causes such as the influence of media exposure, social norms and opportunities for action (Ibid., 30-36). Among the personal characteristics Patchen considers knowledge about the issue, values and ideology, and the type of social relationships valued by a person. In between the underlying and the immediate determinants, Patchen locates the individual's "appraisal of the situation", its seriousness, potential solutions and assgignment of responsibility (Ibid., 8-20).

For each of the underlying determinants, Patchen cites a number of academic studies investigating their particular effects on environmentally significant attitudes and behaviors, both political and private. Among the results for the personal characteristics, he finds that better knowledge about climate change and its causes and consequences is related to a greater likelihood to support "hypothetical referenda" about political countermeasures (Ibid., 24). Regarding values, he finds that individuals who place greater importance on biospheric outcomes and societal welfare rather than personal welfare are more likely to engage in environmentally significant behaviors themselves and to support political solution attempts (Ibid., 25-27). Adopting Ronald Inglehart's terminology of post-materialist value change [cf. @inglehart1987], Patchen concludes, based on his review of empirical studies, that both materialist and non-materialist value system can positively influence environmental attitudes, including support for costly policies, depending on the scope of the political projects and the framing of the problem. In contrast, Inglehart himself finds that post-materialists are more likely to support environmental protection and to vote for environmentalist parties, and that this relationship is strongest in western democracies, where post-materialist values are the most widespread [@inglehart1995]. In addition to the aforementioned dimensions, Patchen finds that individuals who place more value on long-term outcomes, are less utilitarian in their moral framework, and place less importance on conventional authority and private entrepreneurism all exhibit a stronger tendency towards politically environmentalist views (Ibid., 27-30).

It is important to note that, while Patchen considers the effects of both knowledge and personal values, his analysis does not go much further back up the causal chain, i.e. he does not explicitly discuss what shapes these variables apart from socio-demographic characteristics. The findings about these can be briefly summarized as follows: Women are found to be more environmentalist than men in general, which Patchen explains through gender variance in altruism, utilitarianism and risk tolerance, but Patchen also cites some evidence that men are more politically environmentalist and have a stronger tendency to support addressing climate change through public policy (Ibid., 30). Regarding age, Patchen arrives at the "suprising" conclusion that older people tend to engage in more environmentally positive behaviors, but he does not make clear whether this is also true for political attitudes, specifically (Ibid., 31). With regard to social class, Patchen identifies a clear positive relationship between a higher level of formal education and environmentalist behaviors, while the effect of higher income is less clear and more reliant on other factors, such as gender or left-right orientation (Ibid., 32).

@drews2016 offer another overview of the state of research ten years after Patchen's publication. In their cross-disciplinary review of 95 scientific articles, they focus mostly on personal orientations and values, public and individual perception of specific policies, and "contextual factors of policy support, such as social trust, norms and participation, economic and political aspects, geography and weather, media coverage, and framing", while completely bracketing socio-demographic characteristics [@drews2016, 2]. Among their main conclusions, they find that a more left-wing political orientation, as well as egalitarian and self-transcending rather than self-enhancing, hierarchical value systems contribute to a higher level of support for climate policies. Like Patchen, they find "key beliefs" about the existence, anthropogenicity and harmful impact, which stem from "objectively assessed knowledge", to be among the key predictors of policy support (Ibid., 14). They also agree with Patchen that a higher level of emotion (e.g. worry, fear) about climate change is related to a higher level of environmentalism. In the context of a specific policy, individuals' perceptions of its effectiveness and fairness play an important role, and policies that are more coercive and expensive are less likely to garner support. Finally, context conditions including political trust, economic situation and geographic vulnerability, and the influence of the media are found to be relevant, "although the evidence covered here is mixed" (Ibid.).

Some recent papers have focused on specific, prominent or controversial climate policies, most often a carbon dioxide tax. In a survey conducted in France, @douenne2020 find that such a measure draws more support from younger people, which appears intuitive considering the mass movements of climate activism among younger people, such as Fridays for Future, and relativizes Patchen's claim that older people are more environmentalist, at least in the more specific context of divisive climate policies. Further, Douenne and Fabre confirm the importance of left-right orientation, finding that the carbon tax enjoyed support from both centrist and left-leaning respondents, while those leaning to the right tend to reject it [@douenne2020, 10]. @maestre-andres2019 have investigated the role of policy design, with progressive cost distribution and clearly directed use of revenue acting positively. In contrast, @vanboven2018 find that in the US, support for climate policies, including carbon taxation, depends strongly on the partisan affiliation of the politicians proposing them. 

In summary, political environmentalism and citizens' attitudes appear to be the result of a complex interplay of internal and external factors. Internally, values, partisan affiliation, left-right orientation, and objective knowledge about the reality of man-made climate change all contribute to a person being more or less likely to support policies designed to tackle climate change. These relationships are further influenced by context conditions, such as citizens' socio-economic position, the economic and ecological situation of their respective countries, and the way in which climate policies are designed and communicated to the populace. Where, then, could personality become relevant in this model? In the following two subsections, I will explore two separate but mutually compatible explanations. Psychologists have investigated the links between personality traits, personal values, environmentally relevant behaviors, finding that certain traits appear to predispose people to care more about the environment and take actions in order to protect it. It is plausible that these same predispositions will also lead such individuals to develop a more positive stance towards political efforts to limit the impacts of climate change, even if this means making concessions regarding economic concerns and their own personal benefit. As shown by @patchen2006 and @inglehart1995, personal values are important determinants of public support for climate policies, and are known to depend partially on personality traits, so they are a likely intermediary variable. Meanwhile, scholars of political psychology have found personality traits to be related to different modes and outcomes of political information processing and attitude formation. It can be expected that personality traits, through these differences, also have an effect on the attitudes citizens form towards the issue of climate change and political efforts to address it.

## 1.2 Personality traits and environmentalism

The relationship between individuals' innate personality and the values guiding their interactions with the world and other human beings has been an important topic of research in social psychology. In this context, personality is seen as an intrinsic quality which "endures over extended periods of time" [@mondak2010 !!WHERE!!], is "rooted in biology" (Ibid., !!WHERE!!), and is composed of different aspects, referred to as *traits*, which are "internally based psychological characteristics that often correspond to adjectives such as shy, kind, mean, outgoing, dominant, and so forth" [@allen2015 1]. Through a variety of methods, most prominently factor analysis, psychologists have often come to the conclusion that these personality traits can be reduced to five main factors, the *Big Five*. These factors are (I) *Extraversion* or *Surgency*, (II) *Agreeableness*, (III) *Conscientiousness*, (IV) *Neuroticism* or *Emotional (In-)Stability*, and (V) *Openness (to experience)*, *Intellect*, or *Culture* [@goldberg1990; @mccrae1987]. They have been frequently replicated as a basic factor structure in trait psychology and exhibit high cross-observer and cross-instrument reliability [@mccrae1987]. Since they are based on lexical descriptions of personality traits and have been developed mostly on the basis of European and North American respondents, doubts have been expressed as to their applicability in non-western cultural and linguistic contexts [@cheung2001, @schmitt2007a]. However, while this is an important discussion, it is no cause of concern for this paper, as the analysis is restricted to a German sample.

These personality traits have been shown to be related to human values. Values are "basic orientations", which "structure the (political) attitudes of individuals 
and serve to justify behavior" [@scherer2019 209]. @schwartz1992 distinguishes between 10 basic value domains, which are further summarized into two pairs of opposed higher-order values: self-enhancement as opposed to self-transcendence, and openness to change as opposed to conservatism [@schwartz1992 45]. Some studies have linked the Big Five personality traits and the Schwartz value dimensions [cf. @hirsh2007 1584], but more research has been done on subsets of the value system. In a survey of Canadian university students, @hirsh2007 have found that environmentalist values and behaviors are related to higher levels of both Agreeableness and Openness, and that these relationships are independent from each other (Ibid., 1589). They argue that Agreeableness predisposes individuals to hold more environmentalist values because it favors an altruistic disposition, in which individuals expand their self-concept to include others. This extended sense of self can then also include the environment, leading to a higher level of "environmental connectedness" (Ibid., 1586). Openness, which includes cognitive flexibility, may exert the same effect, and is also theorized to "enhance [the individual's] experience of nature, thereby increasing their personal valuation of the natural environment" (Ibid., 1590). @hirsh2010 replicated this result by estimating a structural equation model of personality and environmental concern based on a "much larger community sample of German adults" [@hirsh2010 1]. In addition to Agreeableness and Openness, he also finds the factor of Neuroticism to be positively related to environmental concern. Hirsh interprets this result as showing "a more egoistic form of environmental concern" (Ibid., 248), in which neurotic individuals assign more weight to the negative impacts of environmental degradation. This neurotic response to the threat of environmental decline may plausibly be related to the more recently discussed phenomenon of "climate anxiety", referring to some people's experience of both cognitive-emotional and functional impairment due to their worries about climate change [@clayton2020].

## 1.3 Personality in political psychology

## 1.4 Hypotheses

# 2 Empirical analysis

## 2.1 Dataset

## 2.3 Methods

## 2.4 Results

### 2.4.1 Bivariate Correlations

This chapter references table \ref{table:1}.

```{r table 1, child = 'table1.rmd'}
```

### 2.4.2 Multivariate Model

This chapter references table \ref{table:2}.

```{r table 2, child = 'table2.rmd'}
```

### 2.4.3 Discussion

# Conclusion

\newpage

# References

<div id="refs"></div>

\
\noindent
All web sources last accessed on 25 August 2021.

\newpage

# Appendix A: Replication Notes

This paper was written using RMarkdown [@allaire2021]. All statistical computations were performed using R [@rcoreteam2021]. The full source code for the paper, commented R scripts for all included and supplementary computations, and instructions for replication are publicly available as a Git repository at [https://github.com/Seitzal/climatepsych](https://github.com/Seitzal/climatepsych).

# Appendix B: Supplementary Tables

```{r table 3, child = 'table3.rmd'}
```

```{r table 4, child = 'table4.rmd'}
```

```{r table 5, child = 'table5.rmd'}
```
