# Setup workspace
```{r}
rm(list = ls())               # Clear environment
options(scipen = 4)           # Avoid scientific notation
```

# Import packages
```{r, warning = FALSE}
library(foreign)              # SPSS format
library(tidyverse)            # data cleaning, recoding, and plotting
library(scales)               # rescaling
library(Hmisc)                # correlation matrices
library(psych)                # principal component analysis
library(DescTools)            # logistic regression fit statistics
source("customplots.r")       # correlation heatmaps
```

# Load GLES 2016-2020 panel data
```{r, warning = FALSE}
panel.w1 <- read.spss("ZA6838_w1to9_sA_v4-0-0.sav", to.data.frame = TRUE)   # Waves 1-9 dataset
panel.w12 <- read.spss("ZA6838_w12_sA_v4-0-0.sav", to.data.frame = TRUE)    # Wave 12 dataset
```

```{r}
panel <- inner_join(panel.w1, panel.w12, by = "lfdn")   # Merge the two datasets
panel.vars <- data.frame(names(panel))                  # Extract list of variable names (for reference / exploration)
```

# Derive variables

## Personality trait scores
```{r}
panel <- panel %>%
  # Turn factors for TIPI items into numerics and make exception codes negative
  mutate(
    kp1_perso_a = as.numeric(kp1_2180a) - 4,
    kp1_perso_b = as.numeric(kp1_2180b) - 4,
    kp1_perso_c = as.numeric(kp1_2180c) - 4,
    kp1_perso_d = as.numeric(kp1_2180d) - 4,
    kp1_perso_e = as.numeric(kp1_2180e) - 4,
    kp1_perso_f = as.numeric(kp1_2180f) - 4,
    kp1_perso_g = as.numeric(kp1_2180g) - 4,
    kp1_perso_h = as.numeric(kp1_2180h) - 4,
    kp1_perso_i = as.numeric(kp1_2180i) - 4,
    kp1_perso_j = as.numeric(kp1_2180j) - 4,
    kp1_perso_k = as.numeric(kp1_2180k) - 4,
    kp1_perso_l = as.numeric(kp1_2180l) - 4,
    kp1_perso_m = as.numeric(kp1_2180m) - 4,
    kp1_perso_n = as.numeric(kp1_2180n) - 4,
    kp1_perso_o = as.numeric(kp1_2180o) - 4) %>%
  # Convert exception codes to NAs
  mutate(
    kp1_perso_a = kp1_perso_a %>% replace(kp1_perso_a < 1, NA),
    kp1_perso_b = kp1_perso_b %>% replace(kp1_perso_b < 1, NA),
    kp1_perso_c = kp1_perso_c %>% replace(kp1_perso_c < 1, NA),
    kp1_perso_d = kp1_perso_d %>% replace(kp1_perso_d < 1, NA),
    kp1_perso_e = kp1_perso_e %>% replace(kp1_perso_e < 1, NA),
    kp1_perso_f = kp1_perso_f %>% replace(kp1_perso_f < 1, NA),
    kp1_perso_g = kp1_perso_g %>% replace(kp1_perso_g < 1, NA),
    kp1_perso_h = kp1_perso_h %>% replace(kp1_perso_h < 1, NA),
    kp1_perso_i = kp1_perso_i %>% replace(kp1_perso_i < 1, NA),
    kp1_perso_j = kp1_perso_b %>% replace(kp1_perso_j < 1, NA),
    kp1_perso_k = kp1_perso_k %>% replace(kp1_perso_k < 1, NA),
    kp1_perso_l = kp1_perso_l %>% replace(kp1_perso_l < 1, NA),
    kp1_perso_m = kp1_perso_b %>% replace(kp1_perso_m < 1, NA),
    kp1_perso_n = kp1_perso_n %>% replace(kp1_perso_n < 1, NA),
    kp1_perso_o = kp1_perso_o %>% replace(kp1_perso_o < 1, NA)) %>%
  # Calculate TIPI scale values
  mutate(
    Openness = (kp1_perso_g - kp1_perso_h + 6) / 2,
    Agreeableness = (kp1_perso_n - kp1_perso_o + 6) / 2,
    Extraversion = (kp1_perso_b - kp1_perso_a + 6) / 2,
    Conscientiousness = (kp1_perso_f - kp1_perso_d + 6) / 2,
    Neuroticism = (kp1_perso_k - kp1_perso_l + 6) / 2) %>%
  # Turn factor for risk aversion into numeric and make exception codes negative
  mutate(RiskAversion = as.numeric(kp1_1572) - 4) %>%
  # Convert exception codes to NAs for risk aversion, invert scale and rescale to 1-5
  mutate(RiskAversion = (12 - replace(RiskAversion, RiskAversion < 1, NA)) %>% scales::rescale(c(1,5)))
```

## Climate policy attitudes
```{r}
panel <- panel %>% 
  # Turn factors for policy attitudes into numerics and make exception codes negative
  mutate(
    Priority = as.numeric(kp12_1290) - 17,
    CarbonTax = as.numeric(kp12_2880ak) - 17,
    CombustionBan = as.numeric(kp12_2880w) - 17,
    Nuclear = as.numeric(kp12_2880u) - 17) %>%
  # Convert exception codes into NAs
  mutate(
    Priority = Priority %>% replace(Priority < 1, NA),
    CarbonTax = CarbonTax %>% replace(CarbonTax < 1, NA),
    CombustionBan = CombustionBan %>% replace(CombustionBan < 1, NA),
    Nuclear = Nuclear %>% replace(Nuclear < 1, NA)) %>%
  # Invert priority scale and rescale to 1-5
  mutate(Priority = (8 - Priority) %>% scales::rescale(c(1,5)))
```

## Controls
```{r}
panel <- panel %>%
  # Turn factors for controls into numerics and make exception codes negative
  mutate(
    LeftRight12 = as.numeric(kp12_1500) - 17,
    LeftRight1 = as.numeric(kp1_1500) - 5,
    School9 = as.numeric(kp9_2320) - 7,
    School1 = as.numeric(kp1_2320) - 6,
    SchoolA1 = as.numeric(kpa1_2320) - 6,
    Gender = as.numeric(kpx_2280.x) - 7,
    Age = 2019 - as.numeric(kpx_2290s.x %>% replace(kpx_2290s.x == "1955 und frueher", "1955")),
    Income1 = as.numeric(kp1_2591) - 6,
    IncomeA1 = as.numeric(kpa1_2591) - 6,) %>%
  # Replace exception codes with NA
  mutate(
    LeftRight12 = LeftRight12 %>% replace(LeftRight12 < 1, NA),
    LeftRight1 = LeftRight1 %>% replace(LeftRight1 < 1, NA),
    School9 = School9 %>% replace(School9 < 1 | School9 == 6, NA),
    School1 = School1 %>% replace(School1 < 1 | School1 == 6, NA),
    SchoolA1 = SchoolA1 %>% replace(SchoolA1 < 1 | SchoolA1 == 6, NA),
    Income1 = Income1 %>% replace(Income1 < 1, NA),
    IncomeA1 = IncomeA1 %>% replace(IncomeA1 < 1, NA)) %>%
  # merge responses from different waves, always keeping the newest
  mutate(
    LeftRight = ifelse(is.na(LeftRight12), LeftRight1, LeftRight12),
    Education = ifelse(is.na(School9), ifelse(is.na(School1), SchoolA1, School1), School9),
    Income = ifelse(is.na(IncomeA1), Income1, IncomeA1)) %>%
  # merge school & higher education into one scale
  mutate(Education = Education %>% replace(as.numeric(kp1_2330) == 16 | as.numeric(kpa1_2330) == 16, 6)) 
```

# Shrink dataframe to relevant variables
```{r}
panel.final <- panel %>% select(
  Openness, Agreeableness, Extraversion, Conscientiousness, Neuroticism, RiskAversion,
  Priority, CarbonTax, CombustionBan, Nuclear,
  Age, LeftRight, Gender, Education, Income)
```

# Descriptive and bivariate statistics

## Distribution boxplots

### Personality traits
```{r}
boxplot(panel.final %>% select(Openness, Agreeableness, Extraversion, Conscientiousness, Neuroticism, RiskAversion))
```

### Climate policy attitudes
```{r}
boxplot(panel.final %>% select(Priority, CarbonTax, CombustionBan, Nuclear))
```

## Correlation matrix (Pearson product-moment)
```{r}
rcorr(as.matrix(panel.final), type = "pearson")
plot_corheat(panel.final, method = "pearson", lowerhalf = TRUE)
```

## Correlation matrix (Spearman rank-order)
```{r}
rcorr(as.matrix(panel.final), type = "spearman")
plot_corheat(panel.final, method = "spearman", lowerhalf = TRUE)
```

# Principal component analysis

## Ten-item personality inventory (TIPI): 10 traits, 5 factors
```{r}
panel.t10 <- panel %>% select(
  kp1_perso_a,
  kp1_perso_b,
  #kp1_perso_c,
  kp1_perso_d,
  #kp1_perso_e,
  kp1_perso_f,
  kp1_perso_g,
  kp1_perso_h,
  #kp1_perso_i,
  #kp1_perso_j,
  kp1_perso_k,
  kp1_perso_l,
  #kp1_perso_m,
  kp1_perso_n,
  kp1_perso_o)
pca10 <- principal(panel.t10, nfactors=5, rotate="varimax")
plot_corheat(panel.t10, lowerhalf = TRUE)
summary(pca10)
pca10$loadings
```

## TIPI + 5 additional items: 15 traits
```{r}
panel.t15 <- panel %>% select(
  kp1_perso_a,
  kp1_perso_b,
  kp1_perso_c,
  kp1_perso_d,
  kp1_perso_e,
  kp1_perso_f,
  kp1_perso_g,
  kp1_perso_h,
  kp1_perso_i,
  kp1_perso_j,
  kp1_perso_k,
  kp1_perso_l,
  kp1_perso_m,
  kp1_perso_n,
  kp1_perso_o)
pca15 <- principal(panel.t15, nfactors=5, rotate="varimax")
plot_corheat(panel.t15, lowerhalf = TRUE)
summary(pca15)
pca15$loadings
```

# Regression models

## DV1: Priority

### Pre-filter data
```{r}
panel.final1 <- panel.final %>% select(-c(CarbonTax, CombustionBan, Nuclear)) %>% drop_na()
nrow(panel.final1)
```

### OLS models

#### Controls only
```{r}
mod1.co <- lm(
  data = panel.final1,
  formula = 
    Priority ~ 
    Age + Gender + LeftRight + Education + Income)
summary(mod1.co)
```

#### Full model
```{r}
mod1 <- lm(
  data = panel.final1,
  formula = 
    Priority ~ 
    Openness + Agreeableness + Extraversion + Conscientiousness + Neuroticism + RiskAversion + 
    Age + Gender + LeftRight + Education + Income)
summary(mod1)
```

#### Delta R²
```{r}
summary(mod1)$r.squared - summary(mod1.co)$r.squared
```

#### F-test
```{r}
mod1.anova <- anova(mod1.co, mod1)
mod1.anova
```

### Binary logistic models

#### Controls only
```{r}
mod1.log.co <- glm(
  data = panel.final1 %>% mutate(Priority = as.logical(Priority > 3)),
  family = "binomial",
  formula = Priority ~ 
    Age + Gender + LeftRight + Education + Income)
summary(mod1.log.co)
```

```{r}
mod1.log.co.fits <- PseudoR2(mod1.log.co, which = "all")
mod1.log.co.fits
```

#### Full model
```{r}
mod1.log <- glm(
  data = panel.final1 %>% mutate(Priority = as.logical(Priority > 3)),
  family = "binomial",
  formula = Priority ~ 
    Openness + Agreeableness + Extraversion + Conscientiousness + Neuroticism + RiskAversion + 
    Age + Gender + LeftRight + Education + Income)
summary(mod1.log)
```

```{r}
mod1.log.fits <- PseudoR2(mod1.log, which = "all")
mod1.log.fits
```

#### Delta AIC / BIC
```{r}
mod1.log.fits["AIC"] - mod1.log.co.fits["AIC"]
mod1.log.fits["BIC"] - mod1.log.co.fits["BIC"]
```

#### Chi-square difference test
```{r}
mod1.log.anova <- anova(mod1.log.co, mod1.log, test = "Chisq")
mod1.log.anova
```

## DV2: Carbon tax

### Pre-filter data
```{r}
panel.final2 <- panel.final %>% select(-c(Priority, CombustionBan, Nuclear)) %>% drop_na()
nrow(panel.final2)
```

### OLS models

#### Controls only
```{r}
mod2.co <- lm(
  data = panel.final2,
  formula = 
    CarbonTax ~ 
    Age + Gender + LeftRight + Education + Income)
summary(mod2.co)
```

#### Full model
```{r}
mod2 <- lm(
  data = panel.final2,
  formula = 
    CarbonTax ~ 
    Openness + Agreeableness + Extraversion + Conscientiousness + Neuroticism + RiskAversion + 
    Age + Gender + LeftRight + Education + Income)
summary(mod2)
```

#### Delta R²
```{r}
summary(mod2)$r.squared - summary(mod2.co)$r.squared
```

#### F-test
```{r}
mod2.anova <- anova(mod2.co, mod2)
mod2.anova
```

### Binary logistic models

#### Controls only
```{r}
mod2.log.co <- glm(
  data = panel.final2 %>% mutate(CarbonTax = as.logical(CarbonTax > 3)),
  family = "binomial",
  formula = CarbonTax ~ 
    Age + Gender + LeftRight + Education + Income)
summary(mod2.log.co)
```

```{r}
mod2.log.co.fits <- PseudoR2(mod2.log.co, which = "all")
mod2.log.co.fits
```

#### Full model
```{r}
mod2.log <- glm(
  data = panel.final2 %>% mutate(CarbonTax = as.logical(CarbonTax > 3)),
  family = "binomial",
  formula = CarbonTax ~ 
    Openness + Agreeableness + Extraversion + Conscientiousness + Neuroticism + RiskAversion + 
    Age + Gender + LeftRight + Education + Income)
summary(mod2.log)
```

```{r}
mod2.log.fits <- PseudoR2(mod2.log, which = "all")
mod2.log.fits
```

#### Delta AIC / BIC
```{r}
mod2.log.fits["AIC"] - mod2.log.co.fits["AIC"]
mod2.log.fits["BIC"] - mod2.log.co.fits["BIC"]
```

#### Chi-square difference test
```{r}
mod2.log.anova <- anova(mod2.log.co, mod2.log, test = "Chisq")
mod2.log.anova
```

## DV3: Combustion car ban

### Pre-filter data
```{r}
panel.final3 <- panel.final %>% select(-c(Priority, CarbonTax, Nuclear)) %>% drop_na()
nrow(panel.final3)
```

### OLS models

#### Controls only
```{r}
mod3.co <- lm(
  data = panel.final3,
  formula = 
    CombustionBan ~ 
    Age + Gender + LeftRight + Education + Income)
summary(mod3.co)
```

#### Full model
```{r}
mod3 <- lm(
  data = panel.final3,
  formula = 
    CombustionBan ~ 
    Openness + Agreeableness + Extraversion + Conscientiousness + Neuroticism + RiskAversion + 
    Age + Gender + LeftRight + Education + Income)
summary(mod3)
```

#### Delta R²
```{r}
summary(mod3)$r.squared - summary(mod3.co)$r.squared
```

#### F-test
```{r}
mod3.anova <- anova(mod3.co, mod3)
mod3.anova
```

### Binary logistic models

#### Controls only
```{r}
mod3.log.co <- glm(
  data = panel.final3 %>% mutate(CombustionBan = as.logical(CombustionBan > 3)),
  family = "binomial",
  formula = CombustionBan ~ 
    Age + Gender + LeftRight + Education + Income)
summary(mod3.log.co)
```

```{r}
mod3.log.co.fits <- PseudoR2(mod3.log.co, which = "all")
mod3.log.co.fits
```

#### Full model
```{r}
mod3.log <- glm(
  data = panel.final3 %>% mutate(CombustionBan = as.logical(CombustionBan > 3)),
  family = "binomial",
  formula = CombustionBan ~ 
    Openness + Agreeableness + Extraversion + Conscientiousness + Neuroticism + RiskAversion + 
    Age + Gender + LeftRight + Education + Income)
summary(mod3.log)
```

```{r}
mod3.log.fits <- PseudoR2(mod3.log, which = "all")
mod3.log.fits
```

#### Delta AIC / BIC
```{r}
mod3.log.fits["AIC"] - mod3.log.co.fits["AIC"]
mod3.log.fits["BIC"] - mod3.log.co.fits["BIC"]
```

#### Chi-square difference test
```{r}
mod3.log.anova <- anova(mod3.log.co, mod3.log, test = "Chisq")
mod3.log.anova
```

## DV4: Nuclear power

### Pre-filter data
```{r}
panel.final4 <- panel.final %>% select(-c(Priority, CarbonTax, CombustionBan)) %>% drop_na()
nrow(panel.final4)
```

### OLS models

#### Controls only
```{r}
mod4.co <- lm(
  data = panel.final4,
  formula = 
    Nuclear ~ 
    Age + Gender + LeftRight + Education + Income)
summary(mod4.co)
```

#### Full model
```{r}
mod4 <- lm(
  data = panel.final4,
  formula = 
    Nuclear ~ 
    Openness + Agreeableness + Extraversion + Conscientiousness + Neuroticism + RiskAversion + 
    Age + Gender + LeftRight + Education + Income)
summary(mod4)
```

#### Delta R²
```{r}
summary(mod4)$r.squared - summary(mod4.co)$r.squared
```

#### F-test
```{r}
mod4.anova <- anova(mod4.co, mod4)
mod4.anova
```

### Binary logistic models

#### Controls only
```{r}
mod4.log.co <- glm(
  data = panel.final4 %>% mutate(Nuclear = as.logical(Nuclear > 3)),
  family = "binomial",
  formula = Nuclear ~ 
    Age + Gender + LeftRight + Education + Income)
summary(mod4.log.co)
```

```{r}
mod4.log.co.fits <- PseudoR2(mod4.log.co, which = "all")
mod4.log.co.fits
```

#### Full model
```{r}
mod4.log <- glm(
  data = panel.final4 %>% mutate(Nuclear = as.logical(Nuclear > 3)),
  family = "binomial",
  formula = Nuclear ~ 
    Openness + Agreeableness + Extraversion + Conscientiousness + Neuroticism + RiskAversion + 
    Age + Gender + LeftRight + Education + Income)
summary(mod4.log)
```

```{r}
mod4.log.fits <- PseudoR2(mod4.log, which = "all")
mod4.log.fits
```

#### Delta AIC / BIC
```{r}
mod4.log.fits["AIC"] - mod4.log.co.fits["AIC"]
mod4.log.fits["BIC"] - mod4.log.co.fits["BIC"]
```

#### Chi-square difference test
```{r}
mod4.log.anova <- anova(mod4.log.co, mod4.log, test = "Chisq")
mod4.log.anova
```

# Export results
```{r}
save(file = "results.RData", list = setdiff(ls(), c("panel", "panel.w1", "panel.w12")))
```
